---
title: "Problem Set 4"
author: "Misty Garcia"
date: "25 September 2022"
output: pdf_document
---

Due date: Monday, September 26

For this problem set you will tidy up a dataset of 500 individuals. We also want to calculate each individual's BMI and appropriately categorize them.

Load your data (500_Person_Gender_Height_Weight.csv):

```{r load, include=FALSE}
library(readr)

df <- read_csv('500_Person_Gender_Height_Weight.csv')
df
```

### Question 1

Clean the column headers to be all lower case, have no spaces, and rename "Location information" to location.

```{r clean}
library(tidyverse)

df_clean <- rename_with(df, ~ tolower( gsub(" ", "",.x) ))
df_clean <- rename(df_clean, location=locationinformation)
df_clean
```

\newpage 

### Question 2

Create a new variable that calculates BMI for each individual.

You will need to navigate the different system of measurements (metric vs imperial). Only the United States is using imperial.

* BMI calculation and conversions:
    - metric:   $BMI = weight (kg) / [height (m)]^2$
    - imperial: $BMI = 703 * weight (lbs) / [height (in)]^2$
    - 1 foot = 12 inches
    - 1 cm = 0.01 meter

Although there's many ways you can accomplish this task, we want you to use an if_else() to calculate BMI with the appropriate formula based on each person's location. 

```{r bmi_calc}

distinct(df_clean, location)

#library(dplyr)

df_clean_extend <- df_clean %>% mutate(
  bmi = if_else(location=='New York' | location=='Colorado' | location == 'Hawaii'
          ,703*weight/((height*12)^2)
          ,weight/((height/100)^2)
          ))
df_clean_extend

```

\newpage

### Question 3

Create a new variable that categorizes BMI with case_when():

* Underweight: BMI below 18.5
* Normal: 18.5-24.9
* Overweight: 25.0-29.9
* Obese: 30.0 and Above

```{r bmi_cat}

df_clean_extend <- df_clean_extend %>% mutate( 
  bmi_cat = case_when(
    bmi< 18.5 ~ 'underweight',
    bmi< 25 ~ 'normal',
    bmi< 30 ~ 'overweight',
    TRUE ~ 'obese'
    ))

```

Could we have used if_else()?  

yes, we just would have needed multiple of them

\newpage

### Question 4

Arrange your data first by location and then by descending order of BMI.

```{r arrange_ds}

df_clean_extend_sort <-arrange(df_clean_extend, location, desc(bmi))
df_clean_extend_sort

```

### Question 5

Use a dplyr method to remove the height, weight, and BMI columns from your data.

```{r remove_cols}

df_final <- select(df_clean_extend_sort, -c(height, weight, bmi))
df_final

```

\newpage

### Optional Challenge

Perform all the actions in this problem set with one dpylr call. 

```{r mega_call}

df_test <- read_csv('500_Person_Gender_Height_Weight.csv')

df_test_final <- df_test %>% 
  rename_with(~ tolower( gsub(" ", "",.x) )) %>%
  rename(location=locationinformation) %>%
  mutate(
    bmi = if_else(location=='New York' | location=='Colorado' | location == 'Hawaii'
                  ,703*weight/((height*12)^2)
                  ,weight/((height/100)^2))
    ) %>%
  mutate( 
    bmi_cat = case_when(
      bmi< 18.5 ~ 'underweight',
      bmi< 25 ~ 'normal',
      bmi< 30 ~ 'overweight',
      TRUE ~ 'obese'
      )
    ) %>%
  arrange(location, desc(bmi)) %>%
  select(-c(height, weight, bmi))

df_test_final

```
